Predicting the conformational preferences of proteins using a physics-based free energy method

### 2013-05-09 Paper/paper.tex
May8th_2
9may

First, we validated that our implementation of the confinement method produces results compatible with previous
calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
$\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
in agreement with previous calculations~\cite{Cecchini2009}.

@@

A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
corresponding to better models \cite{Kryshtafovych2011}. Most of the groups use consensus strategies in such experiments.
Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
One of the model was from "PconsR" (GDT-TS = 83, qmode 1 = 0.5865) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
choose these two models as the model by PconsR was the most accurate
predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
free energy using confinement method indicate that the model by PconsR is more stable by $3.9
Kcal/Mol$ which supports
the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
method is able to capture this structure that was otherwise missed.

>>>>

First, we validated that our implementation of the confinement method produces results compatible with previous
calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
$\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
consistent with 200 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
in agreement with previous calculations~\cite{Cecchini2009}.

@@

A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
corresponding to better models \cite{Kryshtafovych2011}. Most of the groups use consensus strategies in such experiments.
Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
One of the model was from "PconsR" (GDT-TS = 96, qmode 1 = 0.5434) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
choose these two models as the model by PconsR was the most accurate
predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
free energy using confinement method indicate that the model by PconsR is more stable by $3.9
Kcal/Mol$ which supports
the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
method is able to capture this structure that was otherwise missed.
